The Dynamics of Anthropomorphic Effects in Human-Robot Interaction

### 2014-03-19 paper.tex
new bib item, new introduction (adapted to the structure: studies - anthropomorphic design - anthropomorphism ...)

The essence of anthropomorphism is the perception of human-like characteristics
in either real or imagined non-human agents. In Human-Robot Interaction (HRI),
anthropomorphism is a commonly observed phenomenon: people tend to ascribe
human-like characteristics to robots, and they are willing to treat and react to
robots as if they were human-social counterparts. The HRI and social robotics
literature is full of various kinds of studies showing how humans readily
interact in a social fashion with robots -- despite the fact that robots are not
humans. In our own studies on HRI analysis in real-world settings, we also
observed anthropomorphism. However, we found qualitative differences and
dynamics over time in how participants anthropomorphized and related to the
robots used in our studies. More concretely, we found that anthropomorphism was
not only based on how much the robot resembles a human (anthropomorphic design)
but that it was determined by multiple factors, including individual
characteristics of the human user, and the context of use, for instance. We also
found that anthropomorphism changed over time with growing experience of the
user with the system. By trying to understand these variances and dynamics in
anthropomorphism, we reviewed related work from the HRI domain and social
robotics. Despite the fact that there is a vast amount of literature on
anthropomorphism in HRI, we were surprised to find that other studies do not
describe similar variances or dynamics. Instead, we came across diverse and
partially conflicting understandings of anthropomorphism. We also had the
impression that the reported data were sometimes over-interpreted (probably due
to the confusion of what exactly anthropomorphism is) or obtained in settings
and assessed in ways that we believe do not give enough consideration to the
complex nature of anthropomorphism. However, when reviewing literature from
social psychology, cognitive science, and neuroscience, we found support for the
variances and dynamics in anthropomorphism that we observed in our studies. As
HRI integrates many disciplines, we think that it is valid to use theories from
these related domains for interpreting our results and that it is relevant to
report on this integration.

>>>>

\subsection*{\hl{Intro when start with anthropomorphic design}}
The Uncanny Valley hypothesis \citep{mori_uncanny_1970} proposes that human-like designs of objects and non-human agents evoke a positive feeling of affinity in the human observer / user, which in turn reflects the experience of emotional engagement with and feelings of empathy for the non-human agent \citep{cheetham_human_2011}. This convenient positive effect of so-called \textbf{anthropomorphic design} is exploited in the design of artificial agents (\eg digital characters, robots) in order to foster people's acceptance of these agents, and to enable effective (human-like and human-social) interaction. 
However, at a certain point (not very well defined, probably due to the various dimensions of human likeness) as the non-human agent approaches but fails to closely resemble a human, the evoked valence in the human user suddenly turns into a negative uncanny feeling of what \cite{mori_uncanny_1970} describes as disgust and revulsion.
The Uncanny Valley hypothesis further predicts that when the non-human agent appears to the human so realistic that it is not distinguishable from a human and in fact perceived to be human, the effect is again a positive one, reflecting feelings and engagement that are similar as those evoked by a human-human interaction.

Besides a feeling of affinity, and empathy, one of the resulting (and  more or less desired) effects of the anthropomorphic design of artificial agents (up to the first positive peak) is the human user's perception of human-likeness in the non-human agent. This phenomenon is labeled as \textbf{anthropomorphism}. There are various reasons and explanations for why people tend to anthropomorphize non-human agents. In general, as anthropomorphism emerges in a (real or imagined) interaction \citep{persson_anthropomorphism_2000} between at least two players (the human who anthropomorphizes and the non-human agent which is anthropomorphized) there are also two main perspectives on the phenomenon: a human-centered (cognitive, psychological) and artifact-centered (design) one. The Uncanny Valley hypothesis can be categorized as being part of latter, artifact-centered approach to understand anthropomorphism.

We investigated anthropomorphism in several human-robot interaction studies that were carried out with real embodied robots, mostly in the real-world.\footnote{Note that the main purpose of these studies was not to analyze in detail the effect of human-likeness in a robot on the user's perception of and interaction with the robot.} We
% now copy-pasted from alternative intro part
found qualitative differences and
dynamics over time in how participants anthropomorphized and related to the
robots used in our studies. More concretely, we found that anthropomorphism was
not only based on how much the robot resembles a human (anthropomorphic design)
but that it was determined by multiple factors, including individual
characteristics of the human user, and the context of use, for instance. We also
found that anthropomorphism changed over time with growing experience of the
user with the system. By trying to understand these variances and dynamics in
anthropomorphism, we reviewed related work from the HRI domain and social
robotics. Despite the fact that there is a vast amount of literature on
anthropomorphism in HRI, we were surprised to find that other studies do not
describe similar variances or dynamics. Instead, we came across diverse and
partially conflicting understandings of anthropomorphism. We also had the
impression that the reported data were sometimes over-interpreted (probably due
to the confusion of what exactly anthropomorphism is) or obtained in settings
and assessed in ways that we believe do not give enough consideration to the
complex nature of anthropomorphism. However, when reviewing literature from
social psychology, cognitive science, and neuroscience, we found support for the
variances and dynamics in anthropomorphism that we observed in our studies. As
HRI integrates many disciplines, we think that it is valid to use theories from
these related domains for interpreting our results and that it is relevant to
report on this integration.

\subsection*{\hl{Intro when start with anthropomorphism}}
The essence of anthropomorphism is the perception of human-like characteristics
in either real or imagined non-human agents. In Human-Robot Interaction (HRI),
anthropomorphism is a commonly observed phenomenon: people tend to ascribe
human-like characteristics to robots, and they are willing to treat and react to
robots as if they were human-social counterparts. The HRI and social robotics
literature is full of various kinds of studies showing how humans readily
interact in a social fashion with robots -- despite the fact that robots are not
humans. In our own studies on HRI analysis in real-world settings, we also
observed anthropomorphism. However, we
%
found qualitative differences and
dynamics over time in how participants anthropomorphized and related to the
robots used in our studies. More concretely, we found that anthropomorphism was
not only based on how much the robot resembles a human (anthropomorphic design)
but that it was determined by multiple factors, including individual
characteristics of the human user, and the context of use, for instance. We also
found that anthropomorphism changed over time with growing experience of the
user with the system. By trying to understand these variances and dynamics in
anthropomorphism, we reviewed related work from the HRI domain and social
robotics. Despite the fact that there is a vast amount of literature on
anthropomorphism in HRI, we were surprised to find that other studies do not
describe similar variances or dynamics. Instead, we came across diverse and
partially conflicting understandings of anthropomorphism. We also had the
impression that the reported data were sometimes over-interpreted (probably due
to the confusion of what exactly anthropomorphism is) or obtained in settings
and assessed in ways that we believe do not give enough consideration to the
complex nature of anthropomorphism. However, when reviewing literature from
social psychology, cognitive science, and neuroscience, we found support for the
variances and dynamics in anthropomorphism that we observed in our studies. As
HRI integrates many disciplines, we think that it is valid to use theories from
these related domains for interpreting our results and that it is relevant to
report on this integration.
